Assessment of Multi-Axial Stress State effects of Circular Concrete-Filled Steel Tubular Columns under Flexural Loading

### 2018-03-28 Paper7.tex
$(data)

\title{Influence of Multi-Axial Stress States on the Flexural Behaviour of Circular Concrete-Filled Steel Tubes}
\author{Yadong Jiang \and António Silva \and Bledar Kalemia \and José Miguel Castro \and Ricardo Monteiro}

@@

	\begin{abstract}
		The paper proposes the methodologies and the equations to consider the influence of multi-axial stress states on the material strength of circular Concrete-filled Steel Tubular (CFST) columns. From the previous experimental and numerical studies on the flexural behaviour of circular CFST members, the combination of multi-axial stress states, namely the concrete confinement effects and the steel bi-axial stress state, was found to benefit the bending capacity of CFST columns. Therefore, the bending capacity of the CFST columns will be underestimated by the Distributed Plasticity (DP) model when uni-axial material properties are used. To account for the material strength variation under multi-axial stress states and to improve the accuracy of the DP model, material strength correction methods were proposed based on the analytical results given by the 3D Finite Element (FE) model. A comprehensive parametric study, which involved a wide range of diameter-thickness ratios, material properties and axial load levels, had been carried out to study the influence of CFST cross-section properties on the multi-axial stress states. Based on the parametric study, the confined concrete strength under bending was found to be independent on the constant axial load level. On the contrary, the steel bi-axial stress was proved to be sensitive to the constant axial load level. A section factor $\xi$ was used to represent both the geometry and material properties of a CFST cross-section. With statistic operations, $\xi$-dependent correction equations were derived to account for the material strength variation under multi-axial stress states. The feasibility of the proposed material strength correction equations was verified to be good for both the design and modelling of circular CFST members.
	\end{abstract}

@@

	The behaviour of concrete-filled steel tubes (CFSTs) is governed by the contribution of the steel tube and the concrete core. As concluded from previous experimental studies (Silva et al. 2016 \cite{Silva2016} and Silva et al. 2017 \cite{RN140}), The concrete core constrains the steel tube to its interior, thus delaying the occurrence of the local buckling to higher levels of deformation and the ductility of the steel tube is improved. Besides its excellent ductility, the composition of concrete and steel will also cause the two components under multi-axial stress state and thus, lead to the material properties variance. The steel tube provides the confined stress to the concrete core and make it under tri-axial stress state, which is well known as concrete confinement effect. The strength and ductility of the concrete core are enhanced by the confinement effect, therefore, the capacity of CFST member under either compression or bending is benefited. In recent years, there were many researchers had investigated the concrete confinement effect of CFST members. Based on the test results, Sakino et al. 2004 \cite{RN41}, Fujimoto et al. 2004 \cite{RN15} and Inai et al. 2004 \cite{RN30} had tested circular and rectangular CFST columns under axial compressive loading, eccentric compressive loading and bending, respectively. All the researchers figured out that, the circular CFST members had a more significant concrete confinement effect than the rectangular columns. Hu et al. 2003 \cite{RN1} and Hu et al. 2005 \cite{RN29} had derived the equations to calculate the confined concrete strength under either compression or bending. The concrete confinement effect of CFST member was found to be sensitive to its diameter-thickness ratio. Regarding Eurocode 4 \cite{RN64}, the concrete confinement effect of circular CFST members is considered in both compressive and bending cases.

@@

	Due to the expansion of the concrete core, the steel tube of CFST will be under bi-axial stress state when the member is loaded. The bi-axial stress state will have an influence on the steel yield stress. According to the research works of Fujimoto et al. 2004 \cite{RN15} and Inai et al. 2004 \cite{RN30}, the bi-axial stress state would increase the steel yield stress when the CFST member is under tension while descrese the steel yield stress when the member is compressed. Regarding the Eurocode 4 \cite{RN64} design provision, it considers a reduction of the steel yield stress when calculating the circular CFST column's plastic resistance to compression. While for the consideration of the flexural resistance of the circular CFST member, the Eurocode 4 is very restrictive since the code does not consider any effects of steel bi-axial stress state to exist.

@@

	The Distributed Plasticity (DP) model is a widely used modelling choice for CFST columns. As the DP model is a simplified numerical solution coupled with fibre element, it could not capture the multi-axial stress states, for both concrete core and steel tube. Therefore, the material properties of CFST should be corrected when applying the DP model. In recent studies of CFST columns, all the researchers (Han 2004 \cite{RN21}, Fujimoto et al. 2004 \cite{RN15}, Inai et al. 2004 \cite{RN30}, Liang et al. 2006 \cite{RN34}, Valipour and Foster 2010 \cite{RN50}, Tort and Hajjar 2010 \cite{RN46}, Zubydan and ElSabbagh 2011 \cite{RN60}, Lai and Varma 2016 \cite{RN32}, etc.) who worked with DP model had considered the concrete strength under confinement effect. But regarding the steel bi-axial stress, only Fujimoto et al. 2004 \cite{RN15}, Inai et al. 2004 \cite{RN30} and Lai and Varma 2016 \cite{RN32} had considered the its influence on the steel yield stress during modelling. The neglecting of steel bi-axial stress states of CFST member may miss leading the analytical results and decrease the accuracy of the DP model. 

@@

	As one may have inferred, there is a clear gap in the consideration of multi-axial stress states, including both concrete confinement effect and steel bi-axial stress state, when model or design the CFST members under flexural loads. The research presented in this paper aims to address this limitation of both the literature and the current design methodologies. This paper starts with a brief description of the multi-axial stress states and the verification of the previous proposed material strength correction method for CFSTs. Based on the 3D Finite Element (FE) model proposed in Jiang et al. 2018 [?], New material strength correction methods, which aims to overcome the aforementioned limitations of the DP model, are developed and validated. By conducting a parametric study on the flexural behaviour of circular CFST members, closed-form expressions which focus on the calculation of the confined concrete stress and the steel bi-axial steel bi-axial yield stress are derived and their feasibility is verified against experimental data.

@@

	When the CFST members are under loading, the composite effects are introduced due to the interaction between concrete core and steel tube. Thus, both the concrete and steel are under the multi-axial stress states. The confined stress provided by the steel tube makes the concrete core under tri-axial stress state, and thus, caused the concrete confinement effect, which can improve its the strength and ductility. Thus, the concrete confinement effect will benefit the flexural capacity of CFST members. There were many researchers studied the concrete confinement effect for CFST members in recent years (Han et al. 2001 \cite{RN23}, Hu et al. 2003 \cite{RN1}, Hu et al. 2005 \cite{RN29}, Tort and Hajjar 2010 \cite{RN46} and Lai and Varma 2016 \cite{RN32}) .

@@

	Not only the concrete core, but also the steel tube is under multi-axial stress state. The concrete core expands under axial load and the hoop stress (or circumferential stress) is generated around the steel tube. As the thickness of the steel tube is relatively small compared to the cross-section diameter, the steel tube could be considered as a shell component. Therefore, the material is under bi-axial stress state (Kwan et al. 2016 \cite{RN31}). Figure \ref{fig-1} shows the stress state of the steel tube of circular CFST member. Since the CFST member is under bending, part of the steel tube is under compression and part of it is under tension and the influences of bi-axial stress state on the two parts are different. As can be seen in Figure \ref{fig-1}, for the compression area, the steel is combined with the compressive stress on vertical direction ($\sigma_2$) and tensile stress on the horizontal direction ($\sigma_1$). While for the tension area, the steel is under tensile stress on both vertical and horizontal directions.

@@

	To consider the steel yield stress variance under bi-axial stress state, the yield surface is introduced (Figure \ref{fig-2}). The positive/negative values on the two axes stand for the tensile/compressive stress, respectively. The hoop stress ($\sigma_h$) is plotted on the \textit{x} axis and the corresponding steel yield stresses under compression ($f_{yc}$) and tension ($f_{yt}$) are annotated on the yield surface. It could be found that, the steel compressive yield stress ($f_{yc}$) is smaller than the uni-axial yield stress ($f_y$), which will have negative influence on the member flexural capacity. On the contrary, the steel tensile yield stress ($f_{yt}$) is larger than the uni-axial yield stress ($f_y$), which will benefit the member strength under bending. For circular CFST member under centrally compression, the whole steel section is under compression. With the influence of bi-axial stress state, the steel yield stress reduces from $f_y$ to $f_{yc}$. For this reason, there is a reduction factor introduced in Eurocode 4 \cite{RN64} to reduce the steel yield stress when considering the plastic resistance of circular CFST to compression. Regarding the flexural capacity of CFST members, the influence of the steel bi-axial stress state on it depends on the relative areas between the compressive/tensile regions as well as the values of $f_{yt}$ and $f_{yc}$.

@@

	To further verify the influence of the steel bi-axial stress state on material strength, the stress versus strain curves of the steel elements located on the same height level are extracted and plotted in Figure \ref{fig-3}. The steel uni-axial stress/strain curve is also plotted in the same figure as the reference curve. The tensile strain/stress are defined as positive values. It could be seen that, the steel compressive yield stress values are less than the uni-axial yield stress and the tensile yield strain is larger than the uni-axial yield stress, which enhances the conclusion obtained from the yield surface.

@@

	In the past, there were many researchers had carried out the studies on the prediction of confined concrete strength for CFST members (Han et al. 2001 \cite{RN23}, Hu et al. 2003 \cite{RN1}, Hu et al. 2005 \cite{RN29} and Tort and Hajjar 2010 \cite{RN46} e.t.c). But in these researches, the influence of steel bi-axial stress state was ignored. Therefore, these derived confined concrete strength equations may result in the inaccuracy of the numerical prediction. Lai and Varma 2016 \cite{RN32} had performed a parametric study on CFST members under compressive loading with the consideration of multi-axial stress states. The equation to calculate the confined concrete strength had been derived. Moreover, the steel compressive yield strength in the research was reduced to $0.9f_y$. But regarding the steel tensile yield stress, there was no correction. Thus, the proposed material strength correction equations of Lai and Varma 2016 \cite{RN32} were suitable for CFST members subjected to axial compression but not sufficient for members under bending.

@@

		\label{tb-1}
	\end{table}
	\par
	Figure \ref{fig-4} displays the lateral load versus drift ratio plots comparison of three results, namely the tests results, the DP model results with uni-axial material strength and the DP model results with corrected material strength. The four plots on the left column of the figure belong show the response of the CFST specimens with no axial load. It could be found that, with the correction methods given by Inai et al. 2004 \cite{RN30} and Fujimoto et al. 2004 \cite{RN15}, the accuracy of the DP model had been improved, as good agreements are observed between the DP model with corrected material strength and the test results. But regarding the specimens with constant axial loads (plots on the right column), nearly no improvement of the DP model results is found by applying the correction methods, by comparing its results with the original DP model predictions. As constant factors were used to correct the steel stress in the correction methods given by the two research works, it could be summarized that the steel bi-axial stress state has a relationship with the axial load level.
	\par
	\begin{figure}[h]
		\centering
		\includegraphics[height=0.9\textheight]{Figures/Figure_5.pdf}
		\caption{Lateral load versus drift ratio plots comparison between the test results, the two DP models}
		\label{fig-4}
	\end{figure}
	\par
	\section{Material Strength Correction Methods}
	As concluded in the Section 2, the material strength correction methods proposed by previous researchers are not efficient enough to improve the prediction results of the DP model, especially in the cases that the circular CFST members are loaded with lateral force combined with constant axial load. Thus, new material strength correction methods should be proposed to account for the influence of the multi-axial stress states on the material strength variance of the circular CFST member. Given that the developed 3D FE model (Jiang et al. 2018 [?]) can accurately represent the flexural behaviour of CFST members, this model is used to calibrate this strength correction methodology. In the following paragraphs, the material strength correction methods, which aims to acquire the confined concrete strength ($f_{cc}$) and the steel bi-axial yield stress ($f_{yb}$) from the numerical results given by the 3D FE model, are detailed.
	\par

@@

	The yield moments of the concrete core and the steel tube are directly related to their material strengths, namely the confined concrete strength and the steel bi-axial yield stress. Thus, to acquire the corrected material strengths, the yield points of the two sections should be addressed first. Figure \ref{fig-6} shows the moment-drift ratio curves of the concrete section. The shape of the moment-drift ratio curve can be classified into two types. For the concrete core with a relatively high confined stress, not only its strength, but also its ductility is remarkably increased. Thus, the corresponding moment-drift ratio curve shows no strength degradation at plastic range, as can be seen in Figure \ref{fig-6} (a). While for the concrete core under low/moderate confinement effect, although its strength and ductility are improved by the confined stress, the confined stress level is not enough to eliminate the influence of the concrete cracks. Thus, there is still strength degradation exhibited in the corresponding moment-drift ratio curve (Figure \ref{fig-6} (b)). For the two conditions, the correction methods are slightly different.

@@

	The method to find the yield point of the concrete section is proposed by ECCS 1986 \cite{RN88}, which used the tangent value to adopt the yield force and displacement of the curve. As marked in Figure \ref{fig-6} (a) and (b), the point ($θ_y^c$, $M_y^c$), which has the tangent value equalling to 10\% of the initial slope of the curve, is adopted as the yield point of the concrete section. Regarding the initial slope, its value equals to the slope of the line which connects the point with $0.4M_{max}^c$ strength to the coordinate origin. For the concrete core with the moment-drift ratio curve type shown in Figure \ref{fig-6} (a), the axial load $N_y^c$ corresponding to drift ratio $θ_y^c$ could be adopted and the axial load and moment pair ($N_y^c$, $M_y^c$) is taken as the section force to acquire the confined concrete strength $f_{cc}$, which is detailed in Section 3.3. Regarding the concrete core with the moment-drift ratio curve type similar to the Figure \ref{fig-6} (b), the influence of the section strength degradation should be considered. The yield moment $M_y^c$ should be compared with the residual moment $M_{res}^c$. If the residual moment $M_{res}^c$ is smaller than the yield moment $M_y^c$, the axial load and moment pair ($N_{res}^c$, $M_{res}^c$) should be taken as the section forces to adopt the confined concrete strength instead of using force pair ($N_y^c$, $M_y^c$). By taking into account the influence of the residual moment $M_{res}^c$, the acquired confined concrete strength should stay at the safe side.

@@

	Figure \ref{fig-7} shows the moment-drift ratio curve of the steel cross-section. As can be seen, the method of adopting the yield point of the steel cross-section is similar to the method applied for the concrete core, which makes use of the tangent value. The moment-drift ratio curve of the steel cross-section shows good ductility due to the constraint effects of the concrete core. The influence of residual strength can be neglected and the force pair ($N_y^s$, $M_y^s$) is used to adopt the steel bi-axial stress, which is described in the following section.

@@

	With the yield points adopted from Section 3.3, the corrected material strengths of the concrete and the steel can be obtained by conducting section analysis. As observed in Jiang et al. 2018 [?], for CFSTs under flexural loading, the stress distributions on the cross-sections are non-uniform. Due to this non-uniformity, it may be difficult to pinpoint the location of the neutral axis of the cross-section. Thus, although the stress distribution on the two sections could be extracted from the numerical results easily, it is not recommended to adopt the corrected material strengths from the non-uniform distributed stress. Therefore, to perform any material strength correction, the non-uniform stress distribution should be “converted” into a uniform distribution of stresses for both sections. Figure \ref{fig-8} shows the simplified stress blocks of the concrete and steel tube sections according to the assumptions made in Section 3.3.1 and 3.3.2. The uniform distributed stress values of on the concrete and the steel cross-sections, which represent the average stress on the two sections, are adopted as the corrected material strengths of the CFST member. In the following subchapters, an indirect approach is proposed to obtain the corrected material strengths of the concrete and steel sections.

@@

		\item The neutral axis is a straight line parallel to the bending axis, i.e. perpendicular to the lateral loading direction. The distance from the neutral axis to the major axis of the section is $h_c$, which ranges from $−0.5D_c$ to $0.5D_c$, as shown in Figure \ref{fig-8}.
	\end{itemize}
	Based on equilibrium of axial loads and moments shown in Equation (\ref{eq-2}) and supplementary Equations (\ref{eq-3}) to (\ref{eq-5}), the confined concrete strength, $f_{cc}$, could be calculated.
	\begin{equation}\label{eq-2}
	\left\{
	\begin{aligned}
	&N_y^c + A_{cc}f_{cc}  & = 0\\
	&M_y^c + A_{cc}f_{cc}h_{cc} & = 0
	\end{aligned}
	\right.
	\textrm{ or }
	\left\{
	\begin{aligned}
	&N_{res}^c + A_{cc}f_{cc} & = 0\\
	&M_{res}^c + A_{cc}f_{cc}h_{cc} & = 0
	\end{aligned}
	\right.
	\end{equation}

@@

		\item The neutral axis is a straight-line parallel to the bending axis. The distance from the neutral axis to the major axis of the section is hs, which ranges from $−0.5D$ to $0.5D$, as shown in Figure \ref{fig-8}.
	\end{itemize}
	\par
	Based on equilibrium of axial loads and moments shown in Equation (6) and the corresponding supplementary Equations (7) to (16), the steel bi-axial yield stress, $f_{yb}$, could be calculated as:

@@

	In order to validate the proposed material strength correction methods for circular CFST members, the eight specimens used in Section 2.2 were applied with the methods. The original and the corrected material strengths are listed in Table \ref{tb-2}. It could be found that, both the concrete and the steel yield strengths are increased by the proposed correction methods, which is in line with the conclusion that the multi-axial stress states will have a positive influences on the flexural capacity of the circular CFST member.

@@

	The eight specimens are re-analysed by the DP model with the material strengths given by the correction methods. As discussed in Section 3.2, the correction methods has already taken the influence of concrete residual strength into consideration during correction procedure. Thus, the acquired confined concrete strength stays at the safe side and there is no need to consider the softening of the concrete strength for a second time. Therefore, the confined concrete constitution law used in DP model is defined as a perfect elasto-plastic curve. The obtained lateral force-drift curves of the eight specimens are shown in Figure \ref{fig-9}. For each member, the numerical results of the DP model with both the uncorrected and corrected material properties (labelled with $FE_{DP}$ and 〖$FE_{DP}^{Method}$, respectively) are plotted against the results of the 3D FE model (labelled with $FE_{3D}$).

@@

	The axial load level will play an important role to the corrected material strength, as the increase of the axial load will result in the change of location of the neutral axis. Thus, five different axial load levels, $μ_n$, are adopted, namely 0\% (no axial load), 5\%, 10\%, 20\% and 30\%. The axial load level is defined as the portion of the plastic axial strength of the cross-section, calculated with Equation (17).

>>>>

\title{Assessment of Multi-Axial Stress State effects of Circular Concrete-Filled Steel Tubular Columns under Flexural Loading}
\author{Yadong Jiang \and António Silva \and Bledar Kalemia \and José Miguel Castro \and Ricardo Monteiro}

@@

	\begin{abstract}
		The paper proposes the methodologies and the equations to consider the influence of multi-axial stress states on the material strength of circular Concrete-filled Steel Tubular (CFST) columns. The combination of multi-axial stress states, namely the concrete confinement effects and the steel bi-axial stress state, is found to benefit the bending capacity of circular CFST columns. Therefore, to account for the material strength variation under multi-axial stress states and to improve the accuracy of the previous proposed Distributed Plasticity (DP) model, material strength correction methods are proposed based on the analytical results given by the 3D Finite Element (FE) model. A comprehensive parametric study, which involved a wide range of diameter-thickness ratios, material properties and axial load levels, has been carried out to study the influence of CFST cross-section properties on the multi-axial stress states. Based on the parametric study, the confined concrete strength under bending is found to be independent of the constant axial load level. On the contrary, the steel bi-axial stress is proved to be sensitive to the cross-section axial load. A section factor $\xi$ is used to represent both the geometry and material properties of a CFST cross-section. With statistic operations, $\xi$-dependent correction equations are derived to account for the material strength variation under multi-axial stress states. The feasibility of the proposed material strength correction equations is verified to be good for both the design and modelling of circular CFST members.
	\end{abstract}

@@

	The behaviour of concrete-filled steel tubes (CFSTs) is governed by the contribution of the steel tube and the concrete core. As concluded from previous experimental studies (Silva et al. 2016 \cite{Silva2016} and Silva et al. 2017 \cite{RN140}), The concrete core constrains the steel tube to its interior, thus delaying the occurrence of the local buckling to higher levels of deformation and the ductility of the steel tube is improved. Besides its excellent ductility, the composition of concrete and steel will also cause the two components under multi-axial stress state and thus, lead to the material properties variance. The steel tube provides the confining stress to the concrete core and makes it under tri-axial stress state, which is well known as concrete confinement effect. The strength and ductility of the concrete core are enhanced by the confinement effect, therefore, the capacity of CFST member under either compression or bending is benefited. In recent years, there were many researchers had investigated the concrete confinement effect of CFST members. Based on the test results, Sakino et al. 2004 \cite{RN41}, Fujimoto et al. 2004 \cite{RN15} and Inai et al. 2004 \cite{RN30} had tested circular and rectangular CFST columns under axial compressive loading, eccentric compressive loading and bending, respectively. All the researchers figured out that, the circular CFST members had a more significant concrete confinement effect than the rectangular columns. Hu et al. 2003 \cite{RN1} and Hu et al. 2005 \cite{RN29} had derived the equations to calculate the confined concrete strength under either compression or bending. The concrete confinement effect of CFST member was found to be sensitive to its diameter-thickness ratio. Regarding Eurocode 4 \cite{RN64}, the concrete confinement effect of circular CFST members is considered in both compressive and bending cases.

@@

	Due to the expansion of the concrete core, the steel tube of CFST will be under bi-axial stress state when the member is loaded. The bi-axial stress state will have an influence on the steel yield stress. According to the research works of Fujimoto et al. 2004 \cite{RN15} and Inai et al. 2004 \cite{RN30}, the bi-axial stress state would increase the steel yield stress when the CFST member is under tension while decrease the steel yield stress when the member is compressed. Regarding the Eurocode 4 \cite{RN64} design provision, it considers a reduction of the steel yield stress when calculating the circular CFST column's plastic resistance to compression. While for the consideration of the flexural resistance of the circular CFST member, the Eurocode 4 is very restrictive since the code does not consider any effects of steel bi-axial stress state to exist.

@@

	The Distributed Plasticity (DP) model is a widely used modelling choice for CFST columns. As the DP model is a simplified numerical solution coupled with fibre element, it could not capture the multi-axial stress states, for both concrete core and steel tube. Therefore, the material properties of CFST should be corrected when applying the DP model. In recent studies of CFST columns, all the researchers (Han 2004 \cite{RN21}, Fujimoto et al. 2004 \cite{RN15}, Inai et al. 2004 \cite{RN30}, Liang et al. 2006 \cite{RN34}, Valipour and Foster 2010 \cite{RN50}, Tort and Hajjar 2010 \cite{RN46}, Zubydan and ElSabbagh 2011 \cite{RN60}, Lai and Varma 2016 \cite{RN32}, etc.) who worked with DP model had considered the concrete strength under confinement effect. But regarding the steel bi-axial stress, only Fujimoto et al. 2004 \cite{RN15}, Inai et al. 2004 \cite{RN30} and Lai and Varma 2016 \cite{RN32} had considered its influence on the steel yield stress during modelling. The neglecting of steel bi-axial stress states of CFST member may miss leading the analytical results and decrease the accuracy of the DP model. 

@@

	As one may have inferred, there is a clear gap in the consideration of multi-axial stress states, including both concrete confinement effect and steel bi-axial stress state, when model or design the CFST members under flexural loads. The research presented in this paper aims to address this limitation of both the literature and the current design methodologies. This paper starts with a brief description of the multi-axial stress states and the verification of the previously proposed material strength correction method for CFSTs. Based on the 3D Finite Element (FE) model proposed in Jiang et al. 2018 [?], New material strength correction methods, which aims to overcome the aforementioned limitations of the DP model, are developed and validated. By conducting a parametric study on the flexural behaviour of circular CFST members, closed-form expressions which focus on the calculation of the confined concrete stress and the steel bi-axial steel bi-axial yield stress are derived and their feasibility is verified against experimental data.

@@

	When the CFST members are under loading, the composite effects are introduced due to the interaction between concrete core and steel tube. Thus, both the concrete and steel are under the multi-axial stress states. The confined stress provided by the steel tube makes the concrete core under tri-axial stress state, and thus, caused the concrete confinement effect, which can improve its the strength and ductility. Thus, the concrete confinement effect will benefit the flexural capacity of CFST members. There were many researchers studied the concrete confinement effect for CFST members in recent years (Han et al. 2001 \cite{RN23}, Hu et al. 2003 \cite{RN1}, Hu et al. 2005 \cite{RN29}, Tort and Hajjar 2010 \cite{RN46} and Lai and Varma 2016 \cite{RN32}).

@@

	Not only the concrete core but also the steel tube is under multi-axial stress state. The concrete core expands under axial load and the hoop stress (or circumferential stress) is generated around the steel tube. As the thickness of the steel tube is relatively small compared to the cross-section diameter, the steel tube could be considered as a shell component. Therefore, the material is under bi-axial stress state (Kwan et al. 2016 \cite{RN31}). Figure \ref{fig-1} shows the stress state of the steel tube of circular CFST member. Since the CFST member is under bending, part of the steel tube is under compression and part of it is under tension and the influences of bi-axial stress state on the two parts are different. As can be seen in Figure \ref{fig-1}, for the compression area, the steel is combined with the compressive stress on the vertical direction ($\sigma_2$) and tensile stress on the horizontal direction ($\sigma_1$). While for the tension area, the steel is under tensile stress on both vertical and horizontal directions.

@@

	To consider the steel yield stress variance under bi-axial stress state, the yield surface is introduced (Figure \ref{fig-2}). The positive/negative values on the two axes stand for the tensile/compressive stress, respectively. The hoop stress ($\sigma_h$) is plotted on the \textit{x} axis and the corresponding steel yield stresses under compression ($f_{yc}$) and tension ($f_{yt}$) are annotated on the yield surface. It could be found that the steel compressive yield stress ($f_{yc}$) is smaller than the uniaxial yield stress ($f_y$), which will have a negative influence on the member flexural capacity. On the contrary, the steel tensile yield stress ($f_{yt}$) is larger than the uniaxial yield stress ($f_y$), which will benefit the member strength under bending. For circular CFST member under central compression, the whole steel section is under compression. With the influence of bi-axial stress state, the steel yield stress reduces from $f_y$ to $f_{yc}$. For this reason, there is a reduction factor introduced in Eurocode 4 \cite{RN64} to reduce the steel yield stress when considering the plastic resistance of circular CFST to compression. Regarding the flexural capacity of CFST members, the influence of the steel bi-axial stress state on it depends on the relative areas between the compressive/tensile regions as well as the values of $f_{yt}$ and $f_{yc}$.

@@

	To further verify the influence of the steel bi-axial stress state on material strength, the stress versus strain curves of the steel elements located on the same height level are extracted and plotted in Figure \ref{fig-3}. The steel uniaxial stress/strain curve is also plotted in the same figure as the reference curve. The tensile strain/stress are defined as positive values. It could be seen that the steel compressive yield stress values are less than the uniaxial yield stress and the tensile yield strain is larger than the uniaxial yield stress, which enhances the conclusion obtained from the yield surface.

@@

	In the past, there were many researchers had carried out the studies on the prediction of confined concrete strength for CFST members (Han et al. 2001 \cite{RN23}, Hu et al. 2003 \cite{RN1}, Hu et al. 2005 \cite{RN29} and Tort and Hajjar 2010 \cite{RN46} e.t.c). But in these research works, the influence of steel bi-axial stress state was ignored. Therefore, these derived confined concrete strength equations may result in the inaccuracy of the numerical prediction. Lai and Varma 2016 \cite{RN32} had performed a parametric study on CFST members under compressive loading with the consideration of multi-axial stress states. The equation to calculate the confined concrete strength had been derived. Moreover, the steel compressive yield strength in the research was reduced to $0.9f_y$. But regarding the steel tensile yield stress, there was no correction. Thus, the proposed material strength correction equations of Lai and Varma 2016 \cite{RN32} were suitable for CFST members subjected to axial compression but not sufficient for members under bending.

@@

		\label{tb-1}
	\end{table}
	\par
	Figure \ref{fig-4} displays the lateral load versus drift ratio plots comparison of three results, namely the tests results, the DP model results with uniaxial material strength and the DP model results with corrected material strength. The four plots on the left column of the figure belong show the response of the CFST specimens with no axial load. It could be found that, with the correction methods given by Inai et al. 2004 \cite{RN30} and Fujimoto et al. 2004 \cite{RN15}, the accuracy of the DP model had been improved, as good agreements are observed between the DP model with corrected material strength and the test results. But regarding the specimens with constant axial loads (plots on the right column), nearly no improvement of the DP model results is found by applying the correction methods, by comparing its results with the original DP model predictions. As constant factors were used to correct the steel stress in the correction methods given by the two research works, it could be summarized that the steel bi-axial stress state has a relationship with the axial load level.
	\par
	\begin{figure}[h]
		\centering
		\includegraphics[height=0.9\textheight]{Figures/Figure_5.pdf}
		\caption{Lateral load versus drift ratio plots comparison between the test results, the two DP models}
		\label{fig-4}
	\end{figure}
	\par
	\section{Material Strength Correction Methods}
	As concluded in Section 2, the material strength correction methods proposed by previous researchers are not efficient enough to improve the prediction results of the DP model, especially in the cases that the circular CFST members are loaded with lateral force combined with a constant axial load. Thus, new material strength correction methods should be proposed to account for the influence of the multi-axial stress states on the material strength variance of the circular CFST member. Given that the developed 3D FE model (Jiang et al. 2018 [?]) can accurately represent the flexural behaviour of CFST members, this model is used to calibrate this strength correction methodology. In the following paragraphs, the material strength correction methods, which aims to acquire the confined concrete strength ($f_{cc}$) and the steel bi-axial yield stress ($f_{yb}$) from the numerical results given by the 3D FE model, are detailed.
	\par

@@

	The yield moments of the concrete core and the steel tube are directly related to their material strengths, namely the confined concrete strength and the steel bi-axial yield stress. Thus, to acquire the corrected material strengths, the yield points of the two sections should be addressed first. Figure \ref{fig-6} shows the moment-drift ratio curves of the concrete section. The shape of the moment-drift ratio curve can be classified into two types. For the concrete core with a relatively high confining stress, not only its strength but also its ductility is remarkably increased. Thus, the corresponding moment-drift ratio curve shows no strength degradation at the plastic range, as can be seen in Figure \ref{fig-6} (a). While for the concrete core under low/moderate confinement effect, although its strength and ductility are improved by the confined stress, the confined stress level is not enough to eliminate the influence of the concrete cracks. Thus, there is still strength degradation exhibited in the corresponding moment-drift ratio curve (Figure \ref{fig-6} (b)). For the two conditions, the correction methods are slightly different.

@@

	The method to find the yield point of the concrete section is proposed by ECCS 1986 \cite{RN88}, which used the tangent value to adopt the yield force and displacement of the curve. As marked in Figure \ref{fig-6} (a) and (b), the point ($θ_y^c$, $M_y^c$), which has the tangent value equalling to 10\% of the initial slope of the curve, is adopted as the yield point of the concrete section. Regarding the initial slope, its value equals to the slope of the line which connects the point with $0.4M_{max}^c$ strength to the coordinate origin. For the concrete core with the moment-drift ratio curve type shown in Figure \ref{fig-6} (a), the axial load $N_y^c$ corresponding to drift ratio $θ_y^c$ could be adopted and the axial load and moment pair ($N_y^c$, $M_y^c$) is taken as the section force to acquire the confined concrete strength $f_{cc}$, which is detailed in Section 3.3. Regarding the concrete core with the moment-drift ratio curve type similar to the Figure \ref{fig-6} (b), the influence of the section strength degradation should be considered. The yield moment $M_y^c$ should be compared with the residual moment $M_{res}^c$. If the residual moment $M_{res}^c$ is smaller than the yield moment $M_y^c$, the axial load and moment pair ($N_{res}^c$, $M_{res}^c$) should be taken as the section forces to adopt the confined concrete strength instead of using force pair ($N_y^c$, $M_y^c$). By taking into account the influence of the residual moment $M_{res}^c$, the acquired confined concrete strength should stay on the safe side.

@@

	Figure \ref{fig-7} shows the moment-drift ratio curve of the steel cross-section. As can be seen, the method of adopting the yield point of the steel cross-section is similar to the method applied to the concrete core, which makes use of the tangent value. The moment-drift ratio curve of the steel cross-section shows good ductility due to the constraint effects of the concrete core. The influence of residual strength can be neglected and the force pair ($N_y^s$, $M_y^s$) is used to adopt the steel bi-axial stress, which is described in the following section.

@@

	With the yield points adopted from Section 3.3, the corrected material strengths of the concrete and the steel can be obtained by conducting section analysis. As observed in Jiang et al. 2018 [?], for CFSTs under flexural loading, the stress distributions on the cross-sections are non-uniform. Due to this non-uniformity, it may be difficult to pinpoint the location of the neutral axis of the cross-section. Thus, although the stress distribution on the two sections could be extracted from the numerical results easily, it is not recommended to adopt the corrected material strengths from the non-uniform distributed stress. Therefore, to perform any material strength correction, the non-uniform stress distribution should be “converted” into a uniform distribution of stresses for both sections. Figure \ref{fig-8} shows the simplified stress blocks of the concrete and steel tube sections according to the assumptions made in Section 3.3.1 and 3.3.2. The uniform distributed stress values on the concrete and the steel cross-sections, which represent the average stress on the two sections, are adopted as the corrected material strengths of the CFST member. In the following subchapters, an indirect approach is proposed to obtain the corrected material strengths of the concrete and steel sections.

@@

		\item The neutral axis is a straight line parallel to the bending axis, i.e. perpendicular to the lateral loading direction. The distance from the neutral axis to the major axis of the section is $h_c$, which ranges from $−0.5D_c$ to $0.5D_c$, as shown in Figure \ref{fig-8}.
	\end{itemize}
	Based on the equilibrium of axial loads and moments shown in Equation (\ref{eq-2}) and supplementary Equations (\ref{eq-3}) to (\ref{eq-5}), the confined concrete strength, $f_{cc}$, could be calculated.
	\begin{equation}\label{eq-2}
	\left\{
	\begin{aligned}
	&N_y^c + A_{cc}f_{cc}  & = 0\\
	&M_y^c + A_{cc}f_{cc}h_{cc} & = 0
	\end{aligned}
	\right.
	\textrm{ or }
	\left\{
	\begin{aligned}
	&N_{res}^c + A_{cc}f_{cc} & = 0\\
	&M_{res}^c + A_{cc}f_{cc}h_{cc} & = 0
	\end{aligned}
	\right.
	\end{equation}

@@

		\item The neutral axis is a straight-line parallel to the bending axis. The distance from the neutral axis to the major axis of the section is hs, which ranges from $−0.5D$ to $0.5D$, as shown in Figure \ref{fig-8}.
	\end{itemize}
	\par
	Based on the equilibrium of axial loads and moments shown in Equation (6) and the corresponding supplementary Equations (7) to (16), the steel bi-axial yield stress, $f_{yb}$, could be calculated as:

@@

	In order to validate the proposed material strength correction methods for circular CFST members, the eight specimens used in Section 2.2 were applied with the methods. The original and the corrected material strengths are listed in Table \ref{tb-2}. It could be found that both the concrete and the steel yield strengths are increased by the proposed correction methods, which is in line with the conclusion that the multi-axial stress states will have a positive influence on the flexural capacity of the circular CFST member.

@@

	The eight specimens are re-analysed by the DP model with the material strengths given by the correction methods. As discussed in Section 3.2, the correction methods have already taken the influence of concrete residual strength into consideration during correction procedure. Thus, the acquired confined concrete strength stays on the safe side and there is no need to consider the softening of the concrete strength for the second time. Therefore, the confined concrete constitution law used in DP model is defined as a perfect elasto-plastic curve. The obtained lateral force-drift curves of the eight specimens are shown in Figure \ref{fig-9}. For each member, the numerical results of the DP model with both the uncorrected and corrected material properties (labelled with $FE_{DP}$ and 〖$FE_{DP}^{Method}$, respectively) are plotted against the results of the 3D FE model (labelled with $FE_{3D}$).

@@

	The axial load level will play an important role in the corrected material strength, as the increase of the axial load will result in the change of location of the neutral axis. Thus, five different axial load levels, $μ_n$, are adopted, namely 0\% (no axial load), 5\%, 10\%, 20\% and 30\%. The axial load level is defined as the portion of the plastic axial strength of the cross-section, calculated with Equation (17).
